---
title: "01-exercises"
author: "Christopher Brown"
date: "April 6, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}
f <- function(x) {
  o = NULL
  for(i in 1:length(x)){
    o = c(o,prod(x[-i]))
  }
  return(o)
}

```
    


## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers except for: 
   multiples of 3 replaced by "Fizz"? 
   multiples of 5 replaced by "Buzz" 
   multiples of 3 and 5 replaced by "FizzBuzz""
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}
f <- function(x) {
  xd3 <- x%%3
  xd3 <- ifelse(xd3==0, "Fizz", "")
  xd5 <- x%%5
  xd5 <- ifelse(xd5==0, "Buzz", "")
  newx <- paste0(xd3,xd5)
  o <- ifelse(x > newx, x, newx)
  return(o)
}
```
