---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Abhishek Chaturvedi"
date: "04/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

library("ggplot2")

myLogistic <- function(x) {
  return(1/(1+exp(-x)))
}

x1 <- c(-10:10)
myLogistic(x1)
ggplot(data.frame(x = x1) , aes(x)) +
  stat_function(fun = myLogistic, geom = "line")


myInverseLogistic <- function(x){
  return(log(x/(1-x)))
}

x2 <- c(0,0.05,0.1,0.2,0.3,0.6,0.8,0.9,0.95,1)
myInverseLogistic(x2)
ggplot(data.frame(x = x2) , aes(x)) +
  stat_function(fun = myInverseLogistic, geom = "line")


```


# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.

```{r "Model for arrival delay more than 22" }

library('data.table')

flightsData <- fread("data/flights.csv")
planesData <- fread("data/planes.csv")
airportsData <- fread("data/airports.csv")
weatherData <- fread("data/weather.csv")

flightsPlanesData <- merge(flightsData,planesData, by.x="tailnum", by.y="tailnum" , all.x = TRUE, suffixes = c(".flights", ".planes"))

flightsPlanesOriginData <- merge(flightsPlanesData,airportsData, by.x = "origin", by.y = "faa", all.x = TRUE, suffixes = c(".flights", ".origin"))

flightsPlanesOriginDestinationData <- merge(flightsPlanesOriginData,airportsData, by.x = "dest", by.y = "faa", all.x = TRUE, suffixes = c(".origin", ".dest"))

flightsPlanesOriginDestinationWeatherData <- merge(flightsPlanesOriginDestinationData,weatherData, by.x = c("year.flights", "month", "day", "origin", "time_hour"), by.y = c("year", "month", "day", "origin", "time_hour"), all.x = TRUE, suffixes = c(".flights", ".weather"), allow.cartesian=FALSE)

flights_data_joined <- flightsPlanesOriginDestinationWeatherData[sample(.N,50000)]

# my old model consists of arr_delay,humid,dep_time,sched_dep_time,sched_arr_time,dep_delay,origin

y <- "arr_delay"
xs <- c('humid','dep_time', 'sched_dep_time','sched_arr_time','dep_delay','origin')

yx <- flights_data_joined[,c(y,xs), with=FALSE]
model1 <- lm(arr_delay ~ ., data=yx)
summary(model1)

# modifying data to create a new column 'gt22' for flights more than 22 hours delay
yx <- within(yx, gt22 <- arr_delay>=22)

model2 <- glm(formula = gt22 ~ . - arr_delay, family=binomial, data=yx)

summary(model2)

# naive model to compare with above model

naiveModel <- glm(formula = gt22 ~ 1, family=binomial, data=yx)

summary(naiveModel)

```

# Question:

Is this a good model?  

AIC has improved compare to the naive model which suggests its a good model


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

## Answer: 
### my old model consists of
arr_delay,humid,dep_time,sched_dep_time,sched_arr_time,dep_delay,origin

out of this humid, dep_delay will not be available so the new model will be

```{r "Model for arrival delay a month in advance" }

y <- "arr_delay"
xs <- c('dep_time', 'sched_dep_time','sched_arr_time','origin')

yx <- flights_data_joined[,c(y,xs), with=FALSE]

model3 <- lm(arr_delay ~ . , data=yx)

summary(model3)

# lets add carrier

y <- "arr_delay"
xs <- c('dep_time', 'sched_dep_time','sched_arr_time','origin','carrier')

yx <- flights_data_joined[,c(y,xs), with=FALSE]
model4 <- lm(arr_delay ~ . , data=yx)

summary(model4)


```

