---
title: "Improving Model Perfromance / Tuning Parameters"
author: "Abhishek Chaturvedi"
date: "`r Sys.Date()`"
output: html_document
---


## Tuning Parameter

Generically and regardless of model type, what are the purposes of a model
tuning parameters?

### Process of adjusting vrious model options to identify the best fit model
#### reducing model bias
#### reducing errors


## Caret Models

This assignment demonstrates the use of caret for constructing models. Each
model should be built and compared using using `Kappa` as the performance
metric calculated using 10-fold repeated cross-validation with 3 folds.

Using the rectangular data that you created for the NYCFlights to create a model
for arr_delay >= 15 minutes.

- glm
- rpart
- knn
- C50
- randomForest
- adaBoost
- Two methods of your choice from the Caret Model List (you will need to install any dependencies)

Save the caret objects with the names provided.

```{r}
# Your work here.
library('data.table')
library('rpart')
library('caret')

flightsDataJoined <- readRDS("flightsDataJoined.rds")
y <- "arr_delay"

# using xs generated from previous exercise
xs <- c('humid','dep_time', 'sched_dep_time','sched_arr_time','dep_delay','origin')
yx <- flightsDataJoined[,c(y,xs),with=FALSE]
yx <- na.omit(yx)
yx <- within(yx, gt15 <- ifelse(arr_delay>=15, "GT15", "LT15"))


set.seed(333)
inTraining <- createDataPartition(yx[,gt15], p = .75, list = FALSE)
trainingData <- yx[inTraining, ]
testingData <- yx[-inTraining, ]

myCtrl <- trainControl(method="repeatedcv",number = 10,repeats = 3)

fit.glm <- train(gt15~humid+dep_time+sched_dep_time+sched_arr_time+dep_delay+origin, data = trainingData, method = "glm", family="binomial", metric="Accuracy", trControl = myCtrl)

fit.knn <- train(gt15~humid+dep_time+sched_dep_time+sched_arr_time+dep_delay+origin, data = trainingData, method = "knn", trControl = myCtrl, tuneLength = 5, metric="Accuracy")

fit.rpart <- train(gt15~humid+dep_time+sched_dep_time+sched_arr_time+dep_delay+origin, data = trainingData, method = "rpart", trControl = myCtrl, tuneLength = 5, metric="Accuracy")

fit.rf <- train(gt15~humid+dep_time+sched_dep_time+sched_arr_time+dep_delay+origin, data = trainingData, method = "rf", trControl = myCtrl, metric="Accuracy")


fit.myown1 <- ..
fit.myown1 <- ..
```

Compare the  models?


```{r}

yhat.fit.glm <- predict(fit.glm,testingData,type="raw")
yhat.fit.knn <- predict(fit.knn,testingData,type="raw")
yhat.fit.rpart <- predict(fit.rpart,testingData,type="raw")
yhat.fit.rf <- predict(fit.rf,testingData,type="raw")

confusionMatrix(data = yhat.fit.glm, reference = testingData$gt15)
confusionMatrix(data = yhat.fit.knn, reference = testingData$gt15)
confusionMatrix(data = yhat.fit.rpart, reference = testingData$gt15)
confusionMatrix(data = yhat.fit.rf, reference = testingData$gt15)


```


Which is best?  Why?

based on confusion matrix rpart is best model
