---
title: "Classification Mettrics"
author: "Abhishek Chaturvedi"
date: "May 17, 2017"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions", echo=FALSE}

# test data
x <- c(1,1,0,1,1,0,1)
x1 <- c(1,1,1,0,0,1,1)

# funciton to get the true positive rate
# function accepts two vectors of observed and predicted values
tpr <- function(yo, yp){
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  return(sum(yo & yp)/sum(yo))
}
tpr(x,x1)

fpr <- function(yo, yp) {
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  return(sum((yp != yo) & yp)/(length(yo)-sum(yo)))
}
fpr(x,x1)

tnr <- function(yo, yp) {
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  return(sum(!yo & !yp)/(length(yo)-sum(yo)))
}
tnr(x,x1)

fnr <- function(yo, yp) {
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  return(sum(!yp & yo)/sum(yo))
}
fnr(x,x1)

# sensitivity is true positive rate
sensitivity <- function (yo, yp){
  return(tpr(yo,yp))
}
sensitivity(x, x1)

#specificity is true negative rate
specificity <- function(yo, yp){
  return(tnr(yo,yp))
}
specificity(x, x1)

# recall is same as true positive rate
recall <- function(yo,yp){
  return(tpr(yo,yp))
}  
recall(x,x1)

# is true positive compare to predictied positives
precision <- function(yo,yp){
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  return(sum(yo & yp)/sum(yp))
}
precision(x,x1)

# number of observed +ve cases in given data set
prevalence <- function (yo){
  if (is.null(yo)) {
    stop("NULL 'yo' value")
  }
  return(sum(yo)/length(yo))
}

prevalence(x1)

# accuracy is total True +ve plus -ve / Observed +ve plus -ve
accuracy <- function(yo, yp){
  if (is.null(yo) || is.null(yp)) {
    stop("NULL 'yo' or 'yp' value")
  }
  if(length(yo) != length(yp)){
    stop(" number of predicted and observed responses are not same ")
  }
  
  return((sum(yo & yp) + sum(!yo & !yp))/length(yo))
}

accuracy(x, x1)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}


```

