---
title: "Assingment NYC Flight Data"
author: "Abhishek Chaturvedi"
date: "April 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}

#install.packages("data.table")

library(data.table)

flightsData <- fread("data/flights.csv")
print(flightsData)

```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}

library(data.table)

flightsData <- fread("data/flights.csv")

hist(flightsData[arr_delay > 0, arr_delay], main="histogram for arrival delays")
hist(flightsData[dep_delay > 0, dep_delay], main="histogram for departure delays")

```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

library(data.table)

flightsData <- fread("data/flights.csv")

# calculating data out of year month and day column and calculating mean arrival delay
arrDelayVsDate <- flightsData[arr_delay > 0, .(mean(arr_delay), as.Date(paste0(year, "/" , month, "/" , day))), by = .(year, month, day)]

plot(arrDelayVsDate[,.(calc_date=V2, delay_arr=V1)], main="Average arrival delays for every date")

```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}

library(data.table)

flightsData <- fread("data/flights.csv")

# generate data using DT
flightsSummary <- flightsData[,.N, by = .(origin, carrier)]

print(flightsSummary)

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}

library(data.table)

flightsData <- fread("data/flights.csv")
planesData <- fread("data/planes.csv")
airportsData <- fread("data/airports.csv")
weatherData <- fread("data/weather.csv")

flightsPlanesData <- merge(flightsData,planesData, by.x="tailnum", by.y="tailnum" , all.x = TRUE, suffixes = c(".flights", ".planes"))

flightsPlanesOriginData <- merge(flightsPlanesData,airportsData, by.x = "origin", by.y = "faa", all.x = TRUE, suffixes = c(".flights", ".origin"))

flightsPlanesOriginDestinationData <- merge(flightsPlanesOriginData,airportsData, by.x = "dest", by.y = "faa", all.x = TRUE, suffixes = c(".origin", ".dest"))

flightsPlanesOriginDestinationWeatherData <- merge(flightsPlanesOriginDestinationData,weatherData, by.x = c("year.flights", "month", "day", "origin", "time_hour"), by.y = c("year", "month", "day", "origin", "time_hour"), all.x = TRUE, suffixes = c(".flights", ".weather"), allow.cartesian=FALSE)

print(flightsPlanesOriginDestinationWeatherData)

```


