---
title: "Caret / Recursive Partitioning"
author: "Abhishek Chaturvedi"
date: "05/15/2017"
output:
  html_document: default
  pdf_document: default
---

```{r init, warning=FALSE, echo=FALSE, message=FALSE}
library(rpart)
library(caret)
# .. Additional libraries
```


## Exercise 1: caret/logistic regression (5 points)

Rebuild your logistic regression model from the previous week, this time using the `caret` package. 

- Calculate the training or apparent performance of the model. 
- Calculate an unbiased measure of performance 
- Create a ROC Curve for your model

Show all work.

```{r logistic model}
# Your Work Here
library('data.table')
flightsDataJoined <- readRDS("flightsDataJoined.rds")

y <- "arr_delay"
xs <- c('humid','dep_time', 'sched_dep_time','sched_arr_time','dep_delay','origin')
yx <- flightsDataJoined[,c(y,xs),with=FALSE]
yx <- na.omit(yx)

set.seed(333)
inTraining <- createDataPartition(na.omit(yx[,arr_delay]), p = .75, list = FALSE)
training <- yx[inTraining, ]
testing <- yx[-inTraining, ]

set.seed(3333)
lmModel1 <- train(arr_delay ~ ., data = training, method = "lm")

library(pROC)

#performance testing using the model
yhat <- predict(lmModel1,testing,type="raw")
postResample(pred = yhat, obs = testing$arr_delay)


```


## Exercise 2: caret/rpart (5 points)

Using the `caret` and `rpart` packages, create a **classification** model for flight delays using your NYC FLight data. Your solution should include:

- The use of `caret` and `rpart` to train a model.
- An articulation of the the problem your are 
- An naive model
- An unbiased calculation of the performance metric
- A plot of your model -- (the actual tree; there are several ways to do this)
- A discussion of your model 



Show and describe all work

```{r rpart model}
# Creating a classification model for flights which has arr_delay more than 15 mins

library("rpart")
library("rpart.plot")

# adding an extea column gt15 to the data with binary values if delay is more than 15 mins
yx <- within(yx, gt15 <- arr_delay>=15)

set.seed(333)
inTraining <- createDataPartition(na.omit(yx[,arr_delay]), p = .75, list = FALSE)
trainingData <- yx[inTraining, ]
testingData <- yx[-inTraining, ]

treeModel <- rpart(gt15 ~ humid + dep_time + sched_dep_time + sched_arr_time + dep_delay + origin, data = trainingData, method = "class")
rpart.plot(treeModel)

treeModel

yhat <- predict(treeModel,testingData,type="class")

confusionMatrix(data = yhat, reference = testingData$gt15)

# the model is depending on dep_delay if its more than 20 the arr_delay will be more than 15 mins have a probability is 0.87

library("pROC")
library("ROCR")

gt15Predicted <- predict(treeModel,testingData,type="vector")

plot(roc(testingData$gt15, gt15Predicted, direction="<"),
     col="yellow", lwd=3, main="flights delayed more than 15 mins")


```





### Questions:

- Discuss the difference between the models and why you would use one model over the other?
- How might you produce an ROC type curve for the *rpart* model? 
