---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "Abhishek Chaturvedi"
date: "04/16/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: Understand Problem 
### Who :
  - Client : Assuming airlines
  - User : Planning department
  - isAffected : Airline staff, customers, airport staff
  
### When :
  - Solution needed ASAP
  - Every Quarter to revist the effectiveness of ETAs
  
### What :
  - Goal(s) Should airline change ETAs for the specific flights, what all factors affect, can airlines predict arrival delays accurately way ahead of time   - ETAs 
  - Data - flights details, origin and destiation weathers and planes all can impact arr delays

### How :
  - using R Studio

## Step 2: Collect and Read Data

```{r "Step 2: Collect and Read Data" }

library(data.table)

flightsData <- fread("data/flights.csv")
planesData <- fread("data/planes.csv")
airportsData <- fread("data/airports.csv")
weatherData <- fread("data/weather.csv")

flightsPlanesData <- merge(flightsData,planesData, by.x="tailnum", by.y="tailnum" , all.x = TRUE, suffixes = c(".flights", ".planes"))

flightsPlanesOriginData <- merge(flightsPlanesData,airportsData, by.x = "origin", by.y = "faa", all.x = TRUE, suffixes = c(".flights", ".origin"))

flightsPlanesOriginDestinationData <- merge(flightsPlanesOriginData,airportsData, by.x = "dest", by.y = "faa", all.x = TRUE, suffixes = c(".origin", ".dest"))

flightsPlanesOriginDestinationWeatherData <- merge(flightsPlanesOriginDestinationData,weatherData, by.x = c("year.flights", "month", "day", "origin", "time_hour"), by.y = c("year", "month", "day", "origin", "time_hour"), all.x = TRUE, suffixes = c(".flights", ".weather"), allow.cartesian=FALSE)

```

## Step 3: Fix Data
- Data has already ben read as data table

## Step 4: Initial Tests

```{r "Naive Model" }
summary(flightsPlanesOriginDestinationWeatherData[,.(arr_delay)])

```

### Identify all the predictors which has strong correlation with arr_delay

```{r "correlation between planes data and arr_delay" }

cor(flightsPlanesOriginDestinationWeatherData[,.(arr_delay,year.planes,engines, seats,speed)],use = "complete.obs")

```

as per above data there is no significant relation between planes data and arr_delay

```{r "correlation between weather and arr_delay" }

cor(flightsPlanesOriginDestinationWeatherData[,.(arr_delay,temp,dewp,humid,wind_dir,wind_speed,wind_gust,precip,pressure,visib)],use = "complete.obs")

```

As per above data arr_delay has a little more relation with humidity

```{r "correlation between airports info and arr_delay" }

cor(flightsPlanesOriginDestinationWeatherData[,.(arr_delay,lat.origin, lat.dest, lon.origin, lon.dest, alt.origin, alt.dest, tz.origin, tz.dest)],use = "complete.obs")

```

No significant relation found between airport geographical details and timezone and arr_delay

```{r "correlation between flights arrival and departure time and arr_delay" }

cor(flightsPlanesOriginDestinationWeatherData[,.(arr_delay,dep_time,sched_dep_time,arr_time,sched_arr_time,dep_delay,air_time,distance)],use = "complete.obs")

```

As per the data arr_delay has strong relation with dep_time, sched_dep_time, sched_arr_time, dep_delay

```{r "correlation between flights year month date time etc and arr_delay" }

cor(flightsPlanesOriginDestinationWeatherData[,.(arr_delay,month,day,hour.flights,minute)],use = "complete.obs")

```

As per the data arr_delay has strong relation with the hour of the day

## Step 5: EDA

```{r "generating model" }

# removing all the data which doesnt seems to have affect on arr_delay or redundent
flightsModelData <- flightsPlanesOriginDestinationWeatherData[,.(arr_delay,humid,dep_time, sched_dep_time,sched_arr_time,dep_delay,origin)]

#generating training data and test data out of the data set
trainingData <- flightsModelData[sample(.N,50000)]
testData1 <- flightsModelData[sample(.N,50000)]
testData2 <- flightsModelData[sample(.N,50000)]
testData3 <- flightsModelData[sample(.N,50000)]

#generating linear model on the seleced predictors
model1 <- lm(arr_delay~., data=trainingData)

summary(predict(model1, testData1))
summary(predict(model1, testData2))
summary(predict(model1, testData3))


```

# Question:

Is this a good model?  (Write your answer here.)

